---
title: "Weight Lifting Exercise - Javier Hernandez Cano"
output: html_notebook
---

First of all, I read this publication [HAR Study](http://groupware.les.inf.puc-rio.br/har). 

Load the data in R into training/testing data:
```{r, message=FALSE, warning=FALSE, error=FALSE}
training <-  read.csv("D:/DatosR/pml-training.csv")
testing <-  read.csv("D:/DatosR/pml-testing.csv")
```


Split the training set into train/test data sets, and to convert the first testing data set into validation data set. The reason is to keep intact the 20 samples that are contained in "pml-testing.csv".

Assign the seed to "12345" to help for the reproducibility:
```{r, message=FALSE, warning=FALSE, error=FALSE}
library(caret)
set.seed(12345)
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
train <- training[inTrain,]
test <- training[-inTrain,]
validation <- testing
```

Explore the target variable:
```{r, message=FALSE, warning=FALSE, error=FALSE}
str(training$classe)
table(training$classe)
table(training$classe)/dim(training)[1]
```

Explore the features:
```{r, message=FALSE, warning=FALSE, error=FALSE}
str(training)
```

Classify the different type of feature (factor, numeric or character):
```{r, message=FALSE, warning=FALSE, error=FALSE}
var_fac <- c()
var_num <- c()
var_char <- c()
for (i in 1:length(train))
{
  if (is.factor(train[,i]))  {var_fac<-append(var_fac,names(train[i]))}
  if (is.numeric(train[,i])) {var_num<-append(var_num,names(train[i]))}
  if (is.character(train[,i])) {var_char<-append(var_char,names(train[i]))}
}
```

Now we are going to put focus in the factors:
```{r}
factors <- train[var_fac]
```

There are a lot of factor features that actually are numeric. So, we have to fix them.  
Select all the wrong factors features (all factors less user_name, cvtd_timestamp, new_window, classe), and tranform them into a numeric features. 
```{r, message=FALSE, warning=FALSE, error=FALSE}
library(dplyr)
factors_wrong <- names(select(factors, -user_name, -cvtd_timestamp, -new_window, -classe))

train2=train[, !(names(train) %in% factors_wrong)]

for (i in factors_wrong) 
  {train2[[i]] <- as.numeric(levels(train[[i]])[train[[i]]])}

train <- train2
#str(train)
```


Now, we analyze the missing values:
```{r, message=FALSE, warning=FALSE, error=FALSE}
library(funModeling)
my_data_status=df_status(train)
#my_data_status[order(-my_data_status$p_na),]
```

Drop features with more than 95% missing values
```{r, message=FALSE, warning=FALSE, error=FALSE}
vars_to_remove_na=subset(my_data_status, my_data_status$p_na > 95)
train=train[, !(names(train) %in% vars_to_remove_na[,"variable"])]
```

Analyze the zero values, and remove features with more than 95% zero values:
```{r, message=FALSE, warning=FALSE, error=FALSE}
vars_to_remove_zero=subset(my_data_status, my_data_status$p_zeros > 95)
train=train[, !(names(train) %in% vars_to_remove_zero[,"variable"])]
```


Let's analyze the remaining missing values:
```{r, message=FALSE, warning=FALSE, error=FALSE}
na_count <-sapply(train, function(y) sum(is.na(y)));na_count
```

we have verified that there are no missing values.

It´s time to modeling:
```{r, message=FALSE, warning=FALSE, error=FALSE, eval=FALSE}
set.seed(4567)

## Parallel Computing
library(doParallel)
cores <- 5
cl <- makeCluster(cores)
registerDoParallel(cores)
getDoParWorkers() # Just checking, how many workers you have


train <- select(train, -X) #drop index variable
control <- trainControl(method = "repeatedcv",number = 10,repeats = 3)
# rf <- train(classe ~ ., data = train, method = "rf", trControl = control)
# saveRDS(rf, file="model_rf.rds")
rf <- readRDS("model_rf.rds")
rf$finalModel
pred <- predict(rf, test)
confusionMatrix(test$classe,pred)
varImp(rf)


train <- select(train, -cvtd_timestamp, -raw_timestamp_part_1,-raw_timestamp_part_2) #drop times variables
# rf2 <- train(classe ~ ., data = train, method = "rf", trControl = control)
# saveRDS(rf2, file="model_rf2.rds")
rf2 <- readRDS("model_rf2.rds")
rf2$finalModel
pred2 <- predict(rf2, test)
confusionMatrix(test$classe,pred2)
varImp(rf2)


train <- select(train, -num_window) 
# rf3 <- train(classe ~ ., data = train, method = "rf", trControl = control)
# saveRDS(rf3, file="model_rf3.rds")
rf3 <- readRDS("model_rf3.rds")
rf3$finalModel
pred3 <- predict(rf3, test)
confusionMatrix(test$classe,pred2)
varImp(rf3)

stopCluster(cl)
```  

I have created a random forest with k-cross valitation (with k=10).  
In three steps, we drop the index feature (X), the times features (cvtd_timestamp, raw_timestamp_part_1,raw_timestamp_part_2) and the num_window feature, because I think without this feature the model will be more generarizable to the validation test. 

The final choosen model is:
```{r, message=FALSE, warning=FALSE, error=FALSE}
rf3
confusionMatrix(test$classe,pred2)
```


And the importance of the variables:
```{r, message=FALSE, warning=FALSE, error=FALSE}
varImp(rf3)
```


Predictions:
```{r, message=FALSE, warning=FALSE, error=FALSE}
(predictions <- predict(rf3,validation))
```
